```{r}
## Load in packages
.libPaths(c("~/R/x86_64-pc-linux-gnu-library/4.4/"));
library(GenomicRanges)
source("src/intervals.R")
```

```{r}
### Unzipped narrowPeak files using the following command: 
# gunzip ~/amc6xa/comp_bio_a08/interval_basics/data/* -k
```


```{r}
### Load in the four datasets as GRange objects 

## Files are TSVs in narrowPeak format with the following 10 columns:
# 1   chrom
# 2   chromStart
# 3   chromEnd
# 4   name        EMPTY in all files
# 5   score       1000 for all rows
# 6   strand      EMPTY (assume +?) in all files
# 7   signalValue 
# 8   pValue      -1 for all rows
# 9   qValue      same float for all rows in the same file
# 10  peak

# Read in TSVs as data.frames (easy to just do manually for just 4 files)
helas3_ctcf_df <- read.table(file = 'data/helas3_ctcf.narrowPeak', sep = '\t', header = F)
helas3_jun_df <- read.table(file = 'data/helas3_jun.narrowPeak', sep = '\t', header = F)
hepg2_ctcf_df <- read.table(file = 'data/hepg2_ctcf.narrowPeak', sep = '\t', header = F)
hepg2_jun_df <- read.table(file = 'data/hepg2_jun.narrowPeak', sep = '\t', header = F)

# Take data.frame columns and create GRanges for each file
helas3_ctcf_gr <- GenomicRanges::GRanges(seqnames=helas3_ctcf_df$V1, ranges=IRanges::IRanges(helas3_ctcf_df$V2, helas3_ctcf_df$V3))
helas3_jun_gr <- GenomicRanges::GRanges(seqnames=helas3_jun_df$V1, ranges=IRanges::IRanges(helas3_jun_df$V2, helas3_jun_df$V3))
hepg2_ctcf_gr <- GenomicRanges::GRanges(seqnames=hepg2_ctcf_df$V1, ranges=IRanges::IRanges(hepg2_ctcf_df$V2, hepg2_ctcf_df$V3))
hepg2_jun_gr <- GenomicRanges::GRanges(seqnames=hepg2_jun_df$V1, ranges=IRanges::IRanges(hepg2_jun_df$V2, hepg2_jun_df$V3))
```

```{r}
gr_list <- list(
  helas3_ctcf = helas3_ctcf_gr,
  helas3_jun = helas3_jun_gr,
  hepg2_ctcf = hepg2_ctcf_gr, 
  hepg2_jun = hepg2_jun_gr
)

pairwise_sims <- pairwiseJaccard(gr_list)
pairwise_sims
```

#### Ranking Similarities
0.607 (helas3_ctcf, hepg2_ctcf)   different cells, same TFs

0.165 (helas3_jun, hepg2_jun)     different cells, same TFs

0.026 (hepg2_ctcf, hepg2_jun)     same cell, different TFs

0.023 (helas3_ctcf, hepg2_jun)    different cells, different TFs

0.020 (helas3_ctcf, helas3_jun)   same cell, different TFs

0.013 (helas3_jun, hepg2_ctcf)    different cells, different TFs


# QUESTIONS
**1. Which two interval sets are the most similar?** (10 pts)
The most similar interval sets by far are the two sets containing CTCF binding sites in HepG2 cells and CTCF binding sites in HeLaS3 cells, with 60.7% of CTCF binding sites across the two cell types being found in both cell types. 

**2. Which two interval sets are the most different?** (10 pts)
The set of binding sites most different are the set of Jun binding sites in HeLaS3 cells and CTCF binding sites in HepG2 cells, with only 1.3% of binding sites found across the two sets being in both sets. 

**3. Based on these results, which factor, CTCF or Jun, would you predict varies more across cell types?** (10 pts) 
Between HeLaS3 and HepG2 cells, Jun binding sites are much more variable than CTCF binding sites. Given that information, I would predict that Jun binding sites are generally more variable between cell types than are CTCF binding sites. 

I think that CTCF binding sites being less variable cross cell types, given CTCFs role in chromatin loop formation, may indicate that chromatin interctions are more consistent across cell types than is Jun activity. At least in this context, Jun might play a larger role in cell-type/state determination than chromatin organization.

**4. Based on these results, do the genomic locations found by ChIP-seq experiments depend more on the cell-type, or on the transcription factor being assayed?** (20 pts)
It seems like transcription factor binding sites (TFBSs) are more similar for the same transcription factors (TFs) in different cell types than they are for different TFs in the same cell type. This result makes sense, since all cell types have the same DNA (with the same motifs), but not all TFs are present in the same concentrations in all cell types, and because chromatin accessibility (which binding motifs are accessible) changes between cell types. 

TFs each bind to somewhat unique motifs in the genome, and thus it makes sense that TFBSs between TFs have little overlap. TFs generally control specific pathways and cellular functions, and it thus makes sense that natural selection would drive the motifs of TFs controlling different cellular functions to be different from one another. 
